Joint Matching and Treatment Effect Estimation in Causal Inference through Bayesian Fellegi-Sunter Methods

### 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
cleaning up repo for submission and in case jerry needs to make edits to tex.

	\begin{figure}[t]
		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}. We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
		\end{center}
	\end{figure}

@@

	\begin{figure}[h!]
		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}. We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
		\end{center}
	\end{figure}

@@

	\begin{figure}[t]
		\begin{center}
			\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}. For each level of overlap and each level of error, we have 100 paired sets of 500 records. Thus this table summarizes results for 900 data files. We see comparable performance for all levels of error and overlap.}
			\label{fig:sadinle_simulation}
		\end{center}
	\end{figure}

@@

\begin{figure}
	\begin{minipage}[c]{0.4\linewidth}
		\begin{center}
		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
		\label{fig:SEI_Z}
		\end{center}
	\end{minipage}
	\hfill
	\begin{minipage}[c]{0.5\linewidth}
		\begin{center}
		\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation.}
		\label{fig:SEI_eval}
	\end{center}
	\end{minipage}%
\end{figure}
	\section{Case Studies}
	\label{sec:case-studies}

@@

	\begin{figure}[t]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
			\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
			\label{fig:overlap-plot}
		\end{center}
	\end{figure}

	\begin{figure}[t]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
			\label{fig:m-and-u}
		\end{center}
	\end{figure}

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
			\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
			\label{fig:nltcs-overlap-plot}
		\end{center}
	\end{figure}

@@

\begin{figure}[t]
	\begin{center}
		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
		\label{fig:sadinle_simulation_partial}
	\end{center}
\end{figure}

@@

	\begin{figure}[!h]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
		\end{center}
	\end{figure}

	\begin{figure}[!h]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
		\end{center}
	\end{figure}

	\begin{figure}[!h]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
		\end{center}
	\end{figure}

@@

	\begin{figure}[h!]
		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$. We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
		\end{center}
	\end{figure}

	\begin{figure}[h!]
		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500. We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
		\end{center}
	\end{figure}

@@

	\begin{figure}[!h]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
		\end{center}
	\end{figure}

	\begin{figure}[!h]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} \label{fig:m_trace}
		\end{center}
	\end{figure}

	\begin{figure}[!h]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} \label{fig:u_trace}
		\end{center}
	\end{figure}
\end{document}

>>>>

	\begin{figure}[t]
		\begin{center}
		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_speed_plot2} 
	% \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}. We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
		\end{center}
	\end{figure}

@@

	\begin{figure}[h!]
		\begin{center} 
		\includegraphics[width=0.7\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides} 
	%	\includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}. We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
		\end{center}
	\end{figure}

@@

	\begin{figure}[t]
		\begin{center}
		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_sim_plot2} 
			%\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}. For each level of overlap and each level of error, we have 100 paired sets of 500 records. Thus this table summarizes results for 900 data files. We see comparable performance for all levels of error and overlap.}
			\label{fig:sadinle_simulation}
		\end{center}
	\end{figure}

@@

\begin{figure}
	\begin{minipage}[c]{0.4\linewidth}
		\begin{center}
		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/Z_plot}
%		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
		\label{fig:SEI_Z}
		\end{center}
	\end{minipage}
	\hfill
	\begin{minipage}[c]{0.5\linewidth}
		\begin{center}
		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/eval_plot} 
%	\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation.}
		\label{fig:SEI_eval}
	\end{center}
	\end{minipage}%
\end{figure}
	\section{Case Studies}
	\label{sec:case-studies}

@@

	\begin{figure}[t]
		\begin{center}
		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_distribution_smallP_bayes}
%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
			\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
			\label{fig:overlap-plot}
		\end{center}
	\end{figure}

	\begin{figure}[t]
		\begin{center}
		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_posterior_smallP} 
%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
			\label{fig:m-and-u}
		\end{center}
	\end{figure}

@@

	\begin{figure}[h]
		\begin{center}
		        \includegraphics[width=0.6\textwidth]{finalFigures/nltcs/overlap_posterior4}
%			\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
			\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
			\label{fig:nltcs-overlap-plot}
		\end{center}
	\end{figure}

@@

\begin{figure}[t]
	\begin{center}
	        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_sim_plot_partial_DR} 
		%\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
		\label{fig:sadinle_simulation_partial}
	\end{center}
\end{figure}

@@

	\begin{figure}[!h]
		\begin{center}
		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_overlap_trace} 
%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
		\end{center}
	\end{figure}

	\begin{figure}[!h]
		\begin{center}
		         \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_m_trace} 
%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
		\end{center}
	\end{figure}

	\begin{figure}[!h]
		\begin{center}
		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_u_trace} 
%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
		\end{center}
	\end{figure}

@@

	\begin{figure}[h!]
		\begin{center} 
		\includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_speed_plot_slides2} 
		%\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$. We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
		\end{center}
	\end{figure}

	\begin{figure}[h!]
		\begin{center}
		 \includegraphics[width=0.6\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides2} 
		% \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500. We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
		\end{center}
	\end{figure}

@@

	\begin{figure}[!h]
		\begin{center}
		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_trace} 
%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
		\end{center}
	\end{figure}

	\begin{figure}[!h]
		\begin{center}
		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_trace} 
%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} 
			\label{fig:m_trace}
		\end{center}
	\end{figure}

	\begin{figure}[!h]
		\begin{center}
		        \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/u_trace} 
%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} 
			\label{fig:u_trace}
		\end{center}
	\end{figure}
\end{document}

### 2023-07-20 presentations/bcm_proposal.tex
cleaning up repo for submission and in case jerry needs to make edits to tex.

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }

\begin{document}
	\title{Bayesian Causal Matching with Fellegi Sunter}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger}
	\institute{Duke University}
	\date{\today}

@@

\begin{frame}{Mahalonobis Distance}
	\begin{itemize}
	\item One method of matching has been to calculate distance in the covariate space through
		\begin{align}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1} |X_1 - X_2|}
		\end{align}
	\item An algorithm then matches each unit in control to closest unit in treated
	\end{itemize}
\end{frame}

\begin{frame}{Propensity Score}
	\begin{itemize}
	\item Another method is to estimate propensity scores $p(Z_n = 1 |X_n)$ for each unit in the data, and match units based on this score.
	\end{itemize}
\end{frame}

\begin{frame}{Pitfalls of Both Approaches}
	\begin{itemize}
		\item The pros and cons of each method are summarized very well on this \href{https://stats.stackexchange.com/questions/511294/what-are-the-pros-and-cons-of-using-mahalanobis-distance-instead-of-propensity-s}{Stack Exchange} page

		\item Essentially, propensity scores can end up matching units with that have similar probability of treatment, but have very different covariate values
		\item Conversely, Mahalonobis can end up matching units that are similar on features that are irrelevant to the outcome of interest
	\end{itemize}
\end{frame}

\begin{frame}{Blending the Two Approaches}
	\begin{itemize}
		\item There is also a good summary on page 5 of \href{https://www.jstatsoft.org/article/view/v042i07}{Sekhot (2011)} recommends including the propensity score into the Malahalonobis distance. This was the genesis for the idea to come
	\end{itemize}
\end{frame}

\section{Proposed Idea}

\begin{frame}{My Idea}
	\begin{itemize}
		\item The machinery of Fellegi Sunter record linkage is very good at identifying duplicate records. It should also be good at identifying separate individuals that are similar.
		\item FS does matching purely on the reliability and uniqueness ($m$ and $u$ parameters) of fields of information. It does not take into account any relationship with an outcome variable. Conceptually, this is very similar to the Mahalonobis distance, but allows for uncertainty quantification.
		\item If we include propensity scores into FS, the model will place high weight on units with ``similar" scores. Then, the FS machinery for the other fields should find the units that are most similar. 
	\end{itemize}
\end{frame}

\begin{frame}{A Similar Idea}
	\begin{itemize}
		\item\href{https://arxiv.org/pdf/2105.02362.pdf}{Alvarez and Levin (2021)} (Section 2, pages 5 - 7) use a Bayesian model to create draws of the propensity scores. For set of scores, they use a deterministic algorithm to match records. 
		\item I imagine that each particular draw of matched pairs may have the issues presented earlier. Propensities are matched, but units may actually look quite different.
		\item Information does not seem to propagate correctly through the Gibbs Sampler
	\end{itemize}
\end{frame}

\begin{frame}{My Idea}
	\begin{itemize}
		\item Data is $\{y_n, z_n, x_n\}_{n = 1}^N$ where $y_n$ is the outcome, $z_n$ is the treatment indicator, and $x_n$ is a vector of $F$ fields of covariates. 
		\item Split data into a datasets $T$ with records $i \in \{1, \ldots, n_T\}$ for treated units and $C$ with records $j \in \{1, \ldots, n_C\}$. 
		\item At each step of the Gibbs Sampler, calculate propensity score $PS(i) = P(Z_i = 1 | X)$ for each data point (logistic or probit regression)
	\end{itemize}
\end{frame}

\begin{frame}{My Idea}
	\begin{itemize}
		\item Define comparison field 
		\begin{align}
			\Gamma_{0} = \begin{cases}
				1 & $PS(i)$ and $PS(j)$ \text{ are sufficently close} \\
				0 & \text{otherwise} \\
			\end{cases}
		\end{align} 
		\item ''Sufficiently close" could mean all units with scores within a certain caliper, or it could mean the $K$ nearest scores.
		\item Conduct Fellegi Sunter Matching on new full set of comparison vectors $\Gamma^{*} = (\Gamma_0, \gamma)$.
		Note $\gamma$ is fixed data, but the $\Gamma$ are random. 
	\end{itemize}
\end{frame}

\begin{frame}
	\begin{itemize}
		\item Each iteration of the Gibbs sampler will provide a matched dataset, from which you can calculate an MCMC sample of the treatment effect (TE) or other quantities of interest
		\item We obtain measures of match quality as usual, through the posterior samples of the linkage structure
		\item I believe this to be first causal matching method to produce matchings probabilistically
		\item The proposed method is fully model based, with no deterministic algorithms. It propogates uncertainty between the matching procedure and the treatment effect estimation. 
	\end{itemize}
\end{frame}

\begin{frame}{Model (Basic)}
	\begin{align}
		Z_n &= x_n' \beta + \epsilon \\
		\beta &\sim N(0, \tau) \\
		\epsilon &\sim N(0, \sigma^2) \\
	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l) \\
		m, u \sim \text{Dirichlet}
		\Delta \sim \text{Desired Record Linkage Prior}
	\end{align}
\end{frame}

\begin{frame}{Model (With Variable Selection)}
	\begin{itemize}
		\item In causal matching, it is especially important not to match on irrelevant features. Therefore, we can use a spike and slab prior on $\beta$, and only use fields where $\beta_f \neq 0$ in the linkage.
	\end{itemize}
\begin{align}
	\beta_j &\sim \pi_0 \delta_0 + \(1 - \pi\)N(0, \tau_j) \\
	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\left[\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l)\right]^{\beta_j \neq 0} \\
\end{align}
\end{frame}

\begin{frame}{Full Conditional for $\beta$}
	\begin{itemize}
		\item The full condition for $\beta_j$ includes a measure of the quality of matches along field $f$
	\end{itemize}
\end{frame}

@@

\begin{frame}{Conclusion}
	\begin{itemize}
		\item Mathematically, any record linkage prior could work. Sadinle, fabl, or original FS (one-to-one, one-to-many, many-to-many) may all be useful , depending on the analysis
		\item If this works, would be a very strong final thesis chapter
	\end{itemize}
\end{frame}

\begin{frame}{Caveats}
	\begin{itemize}
		\item Using a comparison vector made from the propensity score may technically violate the conditional independence assumption of FS
		\item I can't imagine this idea totally failing, but I'm sure that it won't be cutting edge. Sekhot further devised a method that estimates
		\begin{align}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1/2} W S^{-1/2} |X_1 - X_2|}
		\end{align}
		where the entries in $W$ are fine tuned through optimization methods that encourage covariate balance across datasets, based on t-tests and all sorts of other measures. This is almost certainly outperform what I have presented, because it is able to use more information.
		\item Propensity and Mahalonobis matching both have theoretic garuntees. I don't know right now if this method would have them
		\item I know very little about causal inference! This would be a large undertaking
	\end{itemize}
\end{frame}

@@

\begin{frame}{Possible Theoretical Question}
	\begin{itemize}
		\item I'm really struck by the conceptual similarity between Mahalanobis distance and the FS weights. If you make some assumptions (categorical data, one-to-one matching), it might be possible to prove a correspondence between the two. 
	\end{itemize}
\end{frame}

>>>>

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	filecolor=magenta,      
	urlcolor=cyan,
	pdftitle={Overleaf Example},
	pdfpagemode=FullScreen,
}

\begin{document}
	\title{Joint Matching and Treatment Effect Estimation in Causal Inference through Bayesian Fellegi-Sunter Methods}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger}
	\institute{Duke University}
	\date{\today}

@@

\begin{frame}{Mahalonobis Distance and Propensity Matching}
	\begin{itemize}
	\item One approach to matching is to use the Mahalonobis distance, which measures distance between units purely in the covariate space, without regard to which features are related to the outcome of interest. It is defined by
		\begin{align}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1} |X_1 - X_2|}
		\end{align}
	\item Another method is to estimate propensity scores $p(Z_n = 1 |X_n)$ for each unit in the data, and match units based on this score.
	\item Most matching methods in causal inference then use deterministic algorithms to match treated units to their nearest control unit based on the chosen metric
	\end{itemize}
\end{frame}

\begin{frame}{Pitfalls of Both Approaches}
	\begin{itemize}
		\item The pros and cons of each method are summarized very well on this \href{https://stats.stackexchange.com/questions/511294/what-are-the-pros-and-cons-of-using-mahalanobis-distance-instead-of-propensity-s}{Stack Exchange} page
		\item Essentially, propensity scores can end up matching units with that have similar probability of treatment, but have very different covariate values
		\item Conversely, Mahalonobis can end up matching units that are similar on features that are irrelevant to the outcome of interest
	\end{itemize}
\end{frame}

\begin{frame}{Almost Exact Matching (AME)}
	\begin{itemize}
		\item I am new to these readings, but it seems like most of them
		\begin{itemize}
			\item Learn a scoring function (or equivalently, weights on individuals covariates)
			\item Apply scoring function to all units, and find units 
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Additional Thoughts}
	\begin{itemize}
		\item  \href{https://www.jstatsoft.org/article/view/v042i07}{Sekhot (2011)} recommends including the propensity score into the Malahalonobis distance. This is just practical advice, but it forms the foundation of thoughts to come
		\item \href{https://www.lexjansen.com/pharmasug/2006/PublicHealthResearch/PR05.pdf}{This paper} estimates PS, identifies all units in the treatment have PS within a specified caliper, and then chooses the unit with the smallest Mahalanobis distance. No uncertainty quantification
		\item \href{https://arxiv.org/pdf/2105.02362.pdf}{Alvarez and Levin (2021)} (Section 2, pages 5 - 7) use a Bayesian model to create draws of the propensity scores. For set of scores, they use a deterministic algorithm to match records. They do not use any other covariates
	\end{itemize}
\end{frame}

\section{New Research Idea}

\begin{frame}{Intuition}
	\begin{itemize}
		\item The machinery of Fellegi Sunter record linkage is very good at identifying duplicate records. It should also be good at identifying separate individuals that are similar.
		\item FS does matching purely on the reliability and uniqueness ($m$ and $u$ parameters) of fields of information. It does not take into account any relationship with an outcome variable. Conceptually, this is very similar to the Mahalonobis distance, but allows for uncertainty quantification.
		\item If we include propensity scores into FS, the model will place high weight on units with ``similar" scores. Then, the FS machinery for the other fields should find the units that are most similar. 
	\end{itemize}
\end{frame}

\begin{frame}{Nuts and Bolts}
	\begin{itemize}
		\item Data is $\{y_n, z_n, x_n\}_{n = 1}^N$ where $y_n$ is the outcome, $z_n$ is the treatment indicator, and $x_n$ is a vector of $F$ fields of covariates. 
		\item Split data into a datasets $T$ with records $i \in \{1, \ldots, n_T\}$ for treated units and $C$ with records $j \in \{1, \ldots, n_C\}$. 
		\item At each step of the Gibbs Sampler, calculate propensity score $PS(n) = P(Z_n = 1 | X)$ for each data point (logistic or probit regression)
	\end{itemize}
\end{frame}

\begin{frame}{Nuts and Bolts}
	\begin{itemize}
		\item Define comparison field 
		\begin{align*}
			\Gamma_{0} = \begin{cases}
				1 & $PS(i)$ \text{ and } $PS(j)$ \text{ are within specified caliper} \\
				0 & \text{otherwise} \\
			\end{cases}
		\end{align*} 
		\item Conduct Fellegi Sunter Matching on new full set of comparison vectors $\Gamma^{*} = (\Gamma_0, \gamma)$.
		Note $\gamma$ is fixed data, but the $\Gamma$ are random. 
	\end{itemize}
\end{frame}

\begin{frame}{Big Picture}
	\begin{itemize}
		\item Each iteration of the Gibbs sampler will provide a matched dataset, from which you can calculate an MCMC sample of the treatment effect (TE) or other quantities of interest
		\item We obtain measures of match quality as usual, through the posterior samples of the linkage structure
		\item I believe this to be first causal matching method to produce matchings probabilistically
		\item The proposed method is fully model based, with no deterministic algorithms. It propogates uncertainty between the matching procedure and the treatment effect estimation. 
	\end{itemize}
\end{frame}

\section{Proposed Models}

\begin{frame}{Model (Basic)}
	\begin{align*}
		Z_n &= x_n' \beta + \epsilon \\
		\beta &\sim N(0, \tau) \\
		\epsilon &\sim N(0, \sigma^2) \\
	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l)} \\
		m, u &\sim \text{Dirichlet} \\
		\Delta &\sim \text{Desired Record Linkage Prior}
	\end{align*}
Different record linkage priors can induce one-to-one, many-to-one, and many-to-many matches. Any one of them can work within this framework. 
\end{frame}

\begin{frame}{Model (With Variable Selection)}
	\begin{itemize}
		\item In causal matching, it is especially important not to match on irrelevant features. Therefore, we can use a spike and slab prior on $\beta$, and only use fields where $\beta_f \neq 0$ in the linkage.
	\end{itemize}
\begin{align*}
	\beta_j & \sim \pi_0 \delta_0 + (1 - \pi_0)N(0, \tau_j) \\
	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= 
	\prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\left[\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{1 - \Delta_{ij} = 1}\right]^{I(\gamma_{ij}^f = l)}\right]^{\beta_f \neq 0} \\
\end{align*}
\end{frame}

\begin{frame}{Full Conditional for $\beta$}
	Reexpress the likelihood as
		\begin{align*}
			\mathcal{L}(\Delta, m, u, \beta \mid \Gamma^{*}) &= 
			\prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F} c_f \prod_{l=1}^{L_f}\left(  \frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I(\Delta_{ij} = 1)I(\beta_f \neq 0)} \\
			&= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F} c_f \prod_{l=1}^{L_f}w_{fl}^{I(\gamma_{ij}^f = l)I(\Delta_{ij} = 1)I(\beta_f \neq 0)}
		\end{align*}
where $w_{fl} =  \frac{m_{fl}}{u_{fl}}$. Let $M_{fl} = \sum_{i, j} I(\Delta_{ij} = 1)I(\gamma_{ij}^f = l)$.

Then, conditional on $\Delta$, we have 
	\begin{align*}
	\mathcal{L}(m, u, \beta \mid \Gamma^{*}, \Delta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F} \left[c_f \prod_{l=1}^{L_f}w_{fl}^{M_{fl}}\right]^{I(\beta_f \neq 0)}
\end{align*}
\end{frame}

\begin{frame}{Full Conditional for $\beta$}
	The quantity $\prod_{l=1}^{L_f} w_{fl}^{M_{fl}}$ is measure of how strong field $f$ is in the linking procedure. And it shows up in the full conditional for $\beta_f$, showing how the regression and matching procedure are linked. 
\end{frame}

@@

\begin{frame}{Caveats}
	\begin{itemize}
		\item Using a comparison vector made from the propensity score may technically violate the conditional independence assumption of FS
%		\item I can't imagine this idea totally failing, but I'm sure that it won't be cutting edge. Sekhot further devised a method that estimates
%		\begin{align}
%			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1/2} W S^{-1/2} |X_1 - X_2|}
%		\end{align}
%		where the entries in $W$ are fine tuned through optimization methods that encourage covariate balance across datasets, based on t-tests and all sorts of other measures. This is almost certainly outperform what I have presented, because it is able to use more information.
		\item Propensity and Mahalonobis matching both have theoretic garuntees. I don't know right now if this method would have them
		\item I know very little about causal inference! This would be a large undertaking
	\end{itemize}
\end{frame}

@@

\begin{frame}{Possible Theoretical Question}
	\begin{itemize}
		\item I'm really struck by the conceptual similarity between Mahalanobis distance and the FS weights. If you make some assumptions (categorical data, one-to-one matching, known parameters), it might be possible to prove a correspondence between the two. 
	\end{itemize}
\end{frame}

\begin{frame}{Possibly Better Approach}
	\begin{itemize}
		\item The FS model assumes that agreement levels are independent conditional on the matching status of the ``record" pair. This is often violated, with minimal issues.

		However, the propensity score is very much not independent from other agreement levels. It may therefore to incorporate proximity in PS not as a feature, but rather as a \emph{latent class} for the agreement vector.

		See \href{https://projecteuclid.org/journals/annals-of-applied-statistics/volume-13/issue-3/Incorporating-conditional-dependence-in-latent-class-models-for-probabilistic-record/10.1214/19-AOAS1256.full}{Xu et. al (2019)} for more on this.

		I think the first method would work well, and presents enough innovation to stand on its own. I can do this latent class stuff in an extension. 
	\end{itemize}
\end{frame}
